### 2019-05-10 initial/notes.tex
first try at weaktides

Now, let's reconsider the Melnikov integral when perturbing this finite
(non-homoclinic orbit); this may no longer be an exact result but should give
the correct scaling:
\begin{equation}
    M_c = \int\limits_0^T \pd{\phi}{t}\epsilon \p*{1 - \mu^2}\;\mathrm{d}t.
\end{equation}
Naively, we might claim that, since $\pd{\phi}{t}$ changes signs halfway through
the interval of integration, that the only surviving component is $2\epsilon
\bar{\mu}\mu'$, where $\bar{\mu} = \mu_4$ is the average value of $\mu$ and
$\mu'$ is the fluctuation. This gives
\begin{equation}
    M_c = 2\int\limits_0^{2\pi}2\epsilon \frac{\eta \cos I}{1 + \eta \sin I}
        \sqrt{2\eta \sin I \p*{1 - \cos \phi}}\;\mathrm{d}\phi.
\end{equation}
Note that $M_c \propto \eta^{3/2}$, and since the gap opened $\Delta_c =
\frac{M_c}{\pd{\phi}{t}} \propto \eta$, it seems like we're on the right track.
Specifically:
\begin{align}
    M_c &\approx \epsilon 2\eta \cos I A_{sep},\\
    \Delta_c &\approx 2\epsilon \eta \cos I \p*{16 \sqrt{\eta \sin I}}
        \frac{1}{\sqrt{4\eta \sin I}},\\
        &\approx 8\epsilon \eta \cos I.
\end{align}
This also agrees exceedingly well with our simulations This then gives us
hopping probability
\begin{equation}
    P_{hop} = \frac{\Delta_c}{\Delta_M} \approx
        \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.
\end{equation}
This agrees perfectly with the cases we've run.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
        - \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon 2\Omega_1 \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{align}
    \pd{\mu}{(\epsilon\tau)} &\approx \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
        &\approx \p*{1 - \mu^2}
            \p*{\frac{2\Omega_1}{s} - \mu},\\
    \frac{1}{2\Omega_1}\rd{s}{(\epsilon\tau)}
        &\approx \mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2},\nonumber\\
        &\approx - \frac{s}{2\Omega_1}.
\end{align}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$.

>>>>

Now, let's reconsider the Melnikov integral when perturbing this finite
(non-homoclinic orbit); this may no longer be an exact result but should give
the correct scaling:
\begin{equation}
    M_c = \int\limits_0^T \pd{\phi}{t}\epsilon \p*{1 - \mu^2}\;\mathrm{d}t.
\end{equation}
Naively, we might claim that, since $\pd{\phi}{t}$ changes signs halfway through
the interval of integration, that the only surviving component is $2\epsilon
\bar{\mu}\mu'$, where $\bar{\mu} = \mu_4$ is the average value of $\mu$ and
$\mu'$ is the fluctuation. This gives
\begin{equation}
    M_c = 2\int\limits_0^{2\pi}2\epsilon \frac{\eta \cos I}{1 + \eta \sin I}
        \sqrt{2\eta \sin I \p*{1 - \cos \phi}}\;\mathrm{d}\phi.
\end{equation}
Note that $M_c \propto \eta^{3/2}$, and since the gap opened $\Delta_c =
\frac{M_c}{\pd{\phi}{t}} \propto \eta$, it seems like we're on the right track.
Specifically:
\begin{align}
    M_c &\approx \epsilon 2\eta \cos I A_{sep},\\
    \Delta_c &\approx 2\epsilon \eta \cos I \p*{16 \sqrt{\eta \sin I}}
        \frac{1}{\sqrt{4\eta \sin I}},\\
        &\approx 16\epsilon \eta \cos I.
\end{align}
This also agrees exceedingly well with our simulations This then gives us
hopping probability
\begin{equation}
    P_{hop} = \frac{\Delta_c}{\Delta_M} \approx
        \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.
\end{equation}
This agrees perfectly with the cases we've run.

@@

To make the equations more anemable to analytic analysis (not simulation), let's
write down the EOM in $(\mu, \phi)$ coordinates again. The $\phi$ EOM does not
change from the tide-free case, so we can reuse earlier equation:
\begin{align}
    \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
        - \p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi} ,\\
    \pd{\mu}{\tau} &= -\sin I \sin \phi +
        \epsilon \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\\
    \rd{s}{\tau}
        &= \epsilon 2\Omega_1 \p*{\mu - \frac{s}{2\Omega_1}\p*{1 + \mu^2}}.
\end{align}
Assuming $s \gg s_c$ the strong spin-orbit coupling regime, let's first try
assuming $\mu$ is roughly constant over the course of a precession period, then
we can average out the $\phi$ dependencies. Then $\phi$ drops out of the EOM,
and we have approximate averaged equations
\begin{align}
    \pd{\mu}{(\epsilon\tau)} &\approx \frac{2\Omega_1}{s}\p*{1 - \mu^2}
            \p*{1 - \frac{s}{2\Omega_1}\mu},\nonumber\\
        &\approx \p*{1 - \mu^2}
            \p*{\frac{2\Omega_1}{s} - \mu},\\
    \frac{1}{\Omega_1}\rd{s}{(\epsilon\tau)}
        &\approx 2\mu - \frac{s}{\Omega_1}\p*{1 + \mu^2}.
\end{align}
These EOM produce roughly the phase portrait \autoref{fig:phi_avg}.
\begin{figure}[t]
    \centering
    \begin{tikzpicture}
        \draw[<->] (0, 0) -- (8, 0);
        \draw[<->] (0, 0) -- (0, 5);
        \node[below right] at (8, 0) {$s$};
        \node[above left] at (0, 5) {$\mu$};
        \draw (0, 4) -- (2, 4) -- (4, 4);
        \draw[dashed] (4, 4) -- (8, 4);
        \draw (0, 4) -- (2, 4);
        \draw[domain=0:2] plot({\x}, {\x * \x / 2 + 2});
        \node[left] at (0, 4) {$1$};
        \node[left] at (0, 2) {$0$};
        \node[left] at (0, 0) {$-1$};
        \node[below] at (0, 0) {$0$};
        \node[below] at (2, 0) {$\Omega_1$};
        \node[below] at (4, 0) {$2\Omega_1$};
        \draw[domain=4:8] plot({\x}, {8 / (\x - 2)});
        \draw[->] (2, 2) -- (1.5, 2.5);
        \draw[->] (4, 2) -- (3.5, 2.5);
        \draw[->] (6, 0.5) -- (5.5, 1);
        \draw[->] (7, 3.5) -- (6.5, 3);
        \draw[->] (0.5, 3) -- (1, 3.5);
    \end{tikzpicture}
    \caption{Rough phase portrait of $\phi$-averaged equations. Dashed lines
    indicate unstable zeros of at least one of the EOM, while solid lines
    indicate stable zeros of at least one of the EOM\@. The zeros are $\mu = 1$,
    which becomes unstable at $s = 2\Omega_1$, $s = 2\Omega_1/\mu$ and $s' = 0$.
    The only fixed point is $\mu = 1, s = \Omega_1$.}\label{fig:phi_avg}
\end{figure}
In the last term, we note $s \gtrsim 2\Omega_1$ initially, while $\mu \leq 1$,
so we drop both the linear contribution from $\mu$ and approximate $\p*{1 +
\mu^2} \approx 1$ so that $\rd{(s/\Omega_1)}{(\epsilon\tau)} \approx
-s/\Omega_1$.
